---
title: "FE621 assignment3 Problem 4"
author: "Yuxuan Xia"
date: '2018-07-29'
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
We consider 5 stochasitic differential models as follow to fit each columns of 'sample_data.csv' file.

```{r}
sample.data = read.csv("sample_data.csv")
data1 = ts(sample.data[,1],frequency = 365)
data2 = ts(sample.data[,2],frequency = 365)
data3 = ts(sample.data[,3],frequency = 365)
data4 = ts(sample.data[,4],frequency = 365)
data5 = ts(sample.data[,5],frequency = 365)
# Here we read the .csv file to get five cols data
# note here we let frequency = 365, so that dt = 1/365 

fx1 = expression(theta[1]*x)
gx1 = expression(theta[2]*x^(theta[3]))

fx2 = expression(theta[1]+theta[2]*x)
gx2 = expression(theta[3]*x^(theta[4]))

fx3 = expression(theta[1]+theta[2]*x)
gx3 = expression(theta[3]*sqrt(x))

fx4 = expression(theta[1])
gx4 = expression(theta[2]*x^(theta[3]))

fx5 = expression(theta[1]*x)
gx5 = expression(theta[2]+theta[3]*x^(theta[4]))
# we input the expressions of five models
```

## (1) 
In this part, we will use AIC criterion to find out the best model fitting each column data.

### Data 1 (Column 1)
```{r,warning=FALSE,message=FALSE}
library(Sim.DiffProc)

# initial value for all parameters are one
mod1.data1 = fitsde(data = data1,drift = fx1, diffusion = gx1, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod2.data1 = fitsde(data = data1,drift = fx2, diffusion = gx2, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
mod3.data1 = fitsde(data = data1,drift = fx3, diffusion = gx3, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod4.data1 = fitsde(data = data1,drift = fx4, diffusion = gx4, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod5.data1 = fitsde(data = data1,drift = fx5, diffusion = gx5, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')

aic.data1 = c(AIC(mod1.data1),AIC(mod2.data1),AIC(mod3.data1),AIC(mod4.data1),AIC(mod5.data1))
aic.data1
# check the AIC criteria
bestmod.data1 = which(aic.data1==min(aic.data1),arr.ind = TRUE)
bestmod.data1
# shows model 1 fit the data best

```

### Data 2
```{r,warning=FALSE}
mod1.data2 = fitsde(data = data2,drift = fx1, diffusion = gx1, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod2.data2 = fitsde(data = data2,drift = fx2, diffusion = gx2, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
mod3.data2 = fitsde(data = data2,drift = fx3, diffusion = gx3, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod4.data2 = fitsde(data = data2,drift = fx4, diffusion = gx4, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod5.data2 = fitsde(data = data2,drift = fx5, diffusion = gx5, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')

aic.data2 = c(AIC(mod1.data2),AIC(mod2.data2),AIC(mod3.data2),AIC(mod4.data2),AIC(mod5.data2))
aic.data2
# check the AIC criteria
bestmod.data2 = which(aic.data2==min(aic.data2),arr.ind = TRUE)
bestmod.data2
# shows model 1 fit the data best
```

### Data 3 

```{r,warning=FALSE}
mod1.data3 = fitsde(data = data3,drift = fx1, diffusion = gx1, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod2.data3 = fitsde(data = data3,drift = fx2, diffusion = gx2, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
mod3.data3 = fitsde(data = data3,drift = fx3, diffusion = gx3, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod4.data3 = fitsde(data = data3,drift = fx4, diffusion = gx4, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod5.data3 = fitsde(data = data3,drift = fx5, diffusion = gx5, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
aic.data3 = c(AIC(mod1.data3),AIC(mod2.data3),AIC(mod3.data3),AIC(mod4.data3),AIC(mod5.data3))
aic.data3
# check the AIC criteria
bestmod.data3 = which(aic.data3==min(aic.data3),arr.ind = TRUE)
bestmod.data3
```

### Comments
Based on the AIC values, we can find there are three models (model 1,3,4) with the same least aic value. In this situation, we will consider model 1 as the best model fitting data3.

### Data 4 
```{r,warning=FALSE}
mod1.data4 = fitsde(data = data4,drift = fx1, diffusion = gx1, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod2.data4 = fitsde(data = data4,drift = fx2, diffusion = gx2, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
mod3.data4 = fitsde(data = data4,drift = fx3, diffusion = gx3, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod4.data4 = fitsde(data = data4,drift = fx4, diffusion = gx4, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod5.data4 = fitsde(data = data4,drift = fx5, diffusion = gx5, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')

aic.data4 = c(AIC(mod1.data4),AIC(mod2.data4),AIC(mod3.data4),AIC(mod4.data4),AIC(mod5.data4))
aic.data4
# check the AIC criteria
bestmod.data4 = which(aic.data4==min(aic.data4),arr.ind = TRUE)
bestmod.data4
# shows model 5 fit the data best
```

### Data 5
```{r,warning=FALSE}
mod1.data5 = fitsde(data = data5,drift = fx1, diffusion = gx1, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod2.data5 = fitsde(data = data5,drift = fx2, diffusion = gx2, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')
mod3.data5 = fitsde(data = data5,drift = fx3, diffusion = gx3, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod4.data5 = fitsde(data = data5,drift = fx4, diffusion = gx4, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1),pmle = 'euler')
mod5.data5 = fitsde(data = data5,drift = fx5, diffusion = gx5, 
                    start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1),pmle = 'euler')

aic.data5 = c(AIC(mod1.data5),AIC(mod2.data5),AIC(mod3.data5),AIC(mod4.data5),AIC(mod5.data5))
aic.data5
# check the AIC criteria
bestmod.data5 = which(aic.data5==min(aic.data5),arr.ind = TRUE)
bestmod.data5
# shows model 1 fit the data best
```

## (2)

We will implement Euler, Ozaki, Shoji-Ozaki and Kessler method to estimate parameters here.

### Data 1

```{r,warning=FALSE}
pmle = eval(formals(fitsde.default)$pmle)
pmle
# this is the name of Pesudo Maximum likelihood estimator methods
fitres.data1 = lapply(1:4,function(i) fitsde(data = data1, drift = fx1,diffusion = gx1,pmle = pmle[i],
                                      start = list(theta1 = 1,theta2 = 1, theta3 = 1)))
coef.data1 = data.frame(do.call('cbind',lapply(1:4,function(i) coef(fitres.data1[[i]]))))
info.data1 = data.frame(do.call('rbind',lapply(1:4,function(i) logLik(fitres.data1[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) AIC(fitres.data1[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) BIC(fitres.data1[[i]]))),
                        row.names = pmle)
colnames(coef.data1) = pmle
colnames(info.data1) = c('logLik','AIC','BIC')
coef.data1
# coefficient table
info.data1
# information criteria table
best.method.data1 = pmle[which(info.data1$AIC == min(info.data1$AIC),arr.ind = TRUE)]
best.method.data1
# shows the best method is 'Euler'

```

### Data 2

```{r,warning=FALSE}
fitres.data2 = lapply(1:4,function(i) fitsde(data = data2, drift = fx1,diffusion = gx1,pmle = pmle[i],
                                             start = list(theta1 = 1,theta2 = 1, theta3 = 1)))
coef.data2 = data.frame(do.call('cbind',lapply(1:4,function(i) coef(fitres.data2[[i]]))))
info.data2 = data.frame(do.call('rbind',lapply(1:4,function(i) logLik(fitres.data2[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) AIC(fitres.data2[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) BIC(fitres.data2[[i]]))),
                        row.names = pmle)

colnames(coef.data2) = pmle
colnames(info.data2) = c('logLik','AIC','BIC')
coef.data2
# coefficient table
info.data2
# information criteria table
best.method.data2 = pmle[which(info.data2$AIC == min(info.data2$AIC),arr.ind = TRUE)]
best.method.data2
# shows the best method is 'kessler'
```

### Data 3
```{r,warning=FALSE}
fitres.data3 = lapply(1:4,function(i) fitsde(data = data3, drift = fx1,diffusion = gx1,pmle = pmle[i],
                                             start = list(theta1 = 1,theta2 = 1, theta3 = 1)))
coef.data3 = data.frame(do.call('cbind',lapply(1:4,function(i) coef(fitres.data3[[i]]))))
info.data3 = data.frame(do.call('rbind',lapply(1:4,function(i) logLik(fitres.data3[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) AIC(fitres.data3[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) BIC(fitres.data3[[i]]))),
                        row.names = pmle)
colnames(coef.data3) = pmle
colnames(info.data3) = c('logLik','AIC','BIC')
coef.data3
# coefficient table
info.data3
# information criteria table
best.method.data3 = pmle[which(info.data3$AIC == min(info.data3$AIC),arr.ind = TRUE)]
best.method.data3
# we can find 'euler','kessler' and 'ozaki' are the best
```

### Data 4

```{r,warning=FALSE}
fitres.data4 = lapply(1:4,function(i) fitsde(data = data4, drift = fx5,diffusion = gx5,pmle = pmle[i],
                                             start = list(theta1 = 1,theta2 = 1, theta3 = 1,theta4 = 1)))
coef.data4 = data.frame(do.call('cbind',lapply(1:4,function(i) coef(fitres.data4[[i]]))))
info.data4 = data.frame(do.call('rbind',lapply(1:4,function(i) logLik(fitres.data4[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) AIC(fitres.data4[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) BIC(fitres.data4[[i]]))),
                        row.names = pmle)

colnames(coef.data4) = pmle
colnames(info.data4) = c('logLik','AIC','BIC')
coef.data4
# coefficient table
info.data4
# information criteria table
best.method.data4 = pmle[which(info.data4$AIC == min(info.data4$AIC),arr.ind = TRUE)]
best.method.data4
# shows the best method is 'Euler'
```


### Data 5

```{r,warning=FALSE}

fitres.data5 = lapply(1:4,function(i) fitsde(data = data5, drift = fx1,diffusion = gx1,pmle = pmle[i],
                                             start = list(theta1 = 1,theta2 = 1, theta3 = 1)))
coef.data5 = data.frame(do.call('cbind',lapply(1:4,function(i) coef(fitres.data5[[i]]))))
info.data5 = data.frame(do.call('rbind',lapply(1:4,function(i) logLik(fitres.data5[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) AIC(fitres.data5[[i]]))),
                        do.call('rbind',lapply(1:4,function(i) BIC(fitres.data5[[i]]))),
                        row.names = pmle)

colnames(coef.data5) = pmle
colnames(info.data5) = c('logLik','AIC','BIC')
coef.data5
# coefficient table
info.data5
# information criteria table
best.method.data5 = pmle[which(info.data5$AIC == min(info.data5$AIC),arr.ind = TRUE)]
best.method.data5
coef.data5$euler
# shows the best method is 'Euler'
```

### Comments on Result

We will make a table to gather the result of the analysis above.
```{r}
compare.table = data.frame()
compare.table = cbind(c(round(coef.data1$euler,4),NA),c(round(coef.data2$kessler,4),NA),
                      c(round(coef.data3$euler),NA),c(round(coef.data4$euler,4)),
                      c(round(coef.data5$euler,4),NA))
bestmod.col = c(1,1,1,5,1)
# this is the best model for each col data

best.method.compare = cbind('euler','kessler','euler,kessler and ozaki',
                            'euler','euler')
# this is the best method for each col data

compare.table = rbind(bestmod.col,best.method.compare,compare.table)
colnames(compare.table) = c('data1','data2','data3','data4','data5')
row.names(compare.table) = c('model','method','theta1','theta2','theta3','theta4')
compare.table

```

Based on the table above, we can found that Euler method fit most of column of data best(in data 1,3,4 and 5). However, for data 3, we have three method showing the same AIC criteria. Thus, we may not have a good estimation on parameters of data 3.

